Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
purged unification from section 3

\begin{figure}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.7em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenBwdS{n}{d}{r} \; \texttt{+} \; \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
\equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \; \stenCenSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{n}{d}{s}
  \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{irrefl}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.5em]
%
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.5em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
%
%\framebox{${\textit{specDec} :: \texttt{v}} \equiv
%{\textit{specDec'} :: \texttt{v}}$} \\[0.7em]
%\begin{array}{rl}
% MUTUAL
%(\textsc{mutual}) \; &
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; &
%\texttt{stencil} \; \texttt{dependency} (\bar{v})
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w})
%  :: \texttt{v}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
%\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

\begin{example}
Take three array indexing terms
$(i, j+1)$, $(1, j+1)$ and $(3, j)$.
As given by \textsf{schematic}, these have the 
corresponding schemes: $\vtwoh{0}{1}$, $\vtwoh{\infty}{1}$
and $\vtwoh{\infty}{0}$ respectively. The following
relationship of generalises then holds:
%\[
%\xymatrix@R=0.2em@C=2em{
%[\infty \;\; 0] \ar@{-}[dr]^{\preceq_n} &  \\
%& [0 \;\; 1]  \\
%[\infty \;\; 1]  \ar@{-}[ur]_{\preceq_n} & 
%}
%\]
%%
\begin{equation*}
\vtwoh{\infty}{0} \; \not\preceq_2 \; \vtwoh{0}{1} 
\qquad
\vtwoh{\infty}{1} \; \preceq_2 \; \vtwoh{0}{1} 
\end{equation*}
%%
\end{example}

Our terminology of \emph{indexing schemes} and
\emph{unification} are by analogy
with the \emph{type schemes} and \emph{type unification} of
polymorphic types in ML~\cite{milner1978theory}. Our notion is
much more rigid however, there is only one way to unify two
index schemes in our model.

@@

\begin{definition}A \emph{single-entry vector} of size $n$, denoted
$\singleEntry{r}{n}$, is a vector where the $r^{th}$ entry is $1$
and all others are $\infty$, \eg{}, $\singleEntry{1}{2} =
\vtwoh{1}{\infty}$ and $\singleEntry{2}{3} = \vthreeh{\infty}{1}{\infty}$.
Thus, a single-entry vector is an indexing scheme which unifies with
all other schemes which have $1$ as their $r^{th}$ entry,
describing a neighbour offset of $1$. 
%DAO don't understand this sentence:
%In particular the constrained
%requires a non-zero offset from the induction variable.
\end{definition}

@@

Similarly to the single-entry vector, zero-entry vectors 
are also indexing schemes which unify with all other schemes
as long as the $r^{th}$ entry is $0$, representing indices at
the ``origin'' in dimension $r$. 
\end{definition}

@@

The intuition behind $\otimes$ is that it takes
all possible pairs of indexing schemes, treating them as the
lower and upper bounds of an $n$-dimensional rectangle from which the 
remaining vertices of the rectangle are generated (like a bounding
box). \eg{}:
%%
\begin{align*}
\{\vtwoh{1}{2}, \vtwoh{5}{6}\} \otimes \{\vtwoh{3}{4}\} = 
\{&\vtwoh{1}{2}, \vtwoh{3}{2}, \vtwoh{1}{4}, \vtwoh{3}{4}, \\
  & \vtwoh{3}{6}, \vtwoh{5}{4}, \vtwoh{5}{6}\}
\end{align*}
%%
Internally $\otimes$ takes a Cartesian product of the indexing
schemes (all pairs) and then combines each pair using the 
\emph{pairwise permutation} written $\bowtie$ (defined later)
which generates all possible interleavings of two vectors
(\ie{}, all possible non-deterministic choices between $u_i$ and
$v_i$ in each dimension). Some care is then taken over
how this interacts with $\infty$: pariwse permtuation
corresponds to non-determinstic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either 
is $\infty$ then the resultying pairwise permutation is 
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} = 
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
% 
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.
With respect to the notion of unification, the 
%
%
% combines non-$\infty$ constraints
%of the unifiers in different dimensions into less accepting unifiers.
%This means an access pattern unifying with this unifier would require
%both models to be satisfied. Use of $dims$ gives the dimensions that
%constitutent models talk about and is used in the product to eliminate
%general unifier appearing in these dimensions. Allowing general unifier
%in these dimensions would prevent both models to be satisfied
%simultaneously.
%
The notion of which dimensioons $i$ have that \emph{any scheme in the model
has $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
%
\begin{definition}[Constrained dimensions]
Given an $n$-dimensional model $M$, the \emph{constrained dimensions}
of that model is a subset of $\{1, \ldots, n\}$ of 
the dimensions for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}

@@

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\ 
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
% \\
%\setlength{\arraycolsep}{0.5em}
%\left[\begin{array}{ccc}
%0 & 1 & 2 \\
%0 & 1 & 5 \\
%0 & 4 & 2 \\
%0 & 4 & 5 \\
%3 & 1 & 2 \\
%3 & 1 & 5 \\
%3 & 4 & 2 \\
%3 & 4 & 5
%\end{array}
%\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\dnote{I didn't understand mistral's description of unification
wrt to tensor. I can't work it out either! Mistral have another
go given my above explanation}
\end{definition}

>>>>

\begin{figure}
\vspace{-0.2em}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.7em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
 \equiv \; & \stenBwdS{n}{d}{r} \; \texttt{+} \; \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{r \vee s} \\
\equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \; \stenCenSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{r} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{s} \\[0.7em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{r \vee s} \\
\equiv \; & \stenFwdSR{n}{d}{r} \; \texttt{+} \; \stenBwdSR{n}{d}{s}
  \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{R\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{r} \\[0.7em]
%
%
(\textsc{irrefl}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.5em]
%
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.5em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
%
%\framebox{${\textit{specDec} :: \texttt{v}} \equiv
%{\textit{specDec'} :: \texttt{v}}$} \\[0.7em]
%\begin{array}{rl}
% MUTUAL
%(\textsc{mutual}) \; &
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; &
%\texttt{stencil} \; \texttt{dependency} (\bar{v})
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w})
%  :: \texttt{v}
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
%\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

\begin{example}
Take three array indexing terms
$(i, j+1)$, $(1, j+1)$ and $(3, j)$.
As given by \textsf{schematic}, these have the 
corresponding schemes: $\vtwoh{0}{1}$, $\vtwoh{\infty}{1}$
and $\vtwoh{\infty}{0}$ respectively. The following
relationship of generalises then holds:
%\[
%\xymatrix@R=0.2em@C=2em{
%[\infty \;\; 0] \ar@{-}[dr]^{\preceq_n} &  \\
%& [0 \;\; 1]  \\
%[\infty \;\; 1]  \ar@{-}[ur]_{\preceq_n} & 
%}
%\]
%%
\begin{equation*}
\vtwoh{\infty}{0} \; \not\preceq_2 \; \vtwoh{0}{1} 
\qquad
\vtwoh{\infty}{1} \; \preceq_2 \; \vtwoh{0}{1} 
\end{equation*}
%%
\end{example}
\noindent
Our terminology of \emph{indexing schemes} and
\emph{generalisation} are by analogy
with the \emph{type schemes} and \emph{type generalisation} of
polymorphic types in ML~\cite{milner1978theory}. Our notion is
much more rigid however, as our schemes do not bind names
and thus generalisation of does not require substitution; $\infty$ acts
as a positional placeholder.

@@

\begin{definition}A \emph{single-entry vector} of size $n$, denoted
$\singleEntry{r}{n}$, is a vector where the $r^{th}$ entry is $1$
and all others are $\infty$, \eg{}, $\singleEntry{1}{2} =
\vtwoh{1}{\infty}$ and $\singleEntry{2}{3} = \vthreeh{\infty}{1}{\infty}$.
Thus, a single-entry vector is an indexing scheme which generalises 
all other schemes which at least have $1$ as their $r^{th}$ entry,
describing a neighbour offset of $1$. 
%DAO don't understand this sentence:
%In particular the constrained
%requires a non-zero offset from the induction variable.
\end{definition}

@@

Similarly to the single-entry vector, zero-entry vectors 
are also indexing schemes which generalising all other schemes
which at least have the $r^{th}$ entry is $0$, representing indices at
the ``origin'' in dimension $r$. 
\end{definition}

@@

The intuition behind $\otimes$ is that it takes
all possible pairs of indexing schemes, treating them as the
lower and upper bounds of an $n$-dimensional rectangle from which the 
remaining vertices of the rectangle are generated (like a bounding
box). \eg{}:
%%
\begin{align*}
\{\vtwoh{1}{2}, \vtwoh{5}{6}\} \otimes \{\vtwoh{3}{4}\} = 
\{&\vtwoh{1}{2}, \vtwoh{3}{2}, \vtwoh{1}{4}, \vtwoh{3}{4}, \\
  & \vtwoh{3}{6}, \vtwoh{5}{4}, \vtwoh{5}{6}\}
\end{align*}
%%
Internally $\otimes$ takes a Cartesian product of the indexing
schemes (all pairs) and then combines each pair using the 
\emph{pairwise permutation} written $\bowtie$ (defined later)
which generates all possible interleavings of two vectors
(\ie{}, all possible non-deterministic choices between $u_i$ and
$v_i$ in each dimension). Some care is then taken over
how this interacts with $\infty$: pariwse permtuation
corresponds to non-determinstic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either 
is $\infty$ then the resultying pairwise permutation is 
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} = 
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
% 
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.
%With respect to the notion of unification, the 
%
%
% combines non-$\infty$ constraints
%of the unifiers in different dimensions into less accepting unifiers.
%This means an access pattern generalises with this scheme would require
%both models to be satisfied. Use of $dims$ gives the dimensions that
%constitutent models talk about and is used in the product to eliminate
%general unifier appearing in these dimensions. Allowing general unifier
%in these dimensions would prevent both models to be satisfied
%simultaneously.
%  u in M * N   u < w
% mplies that 
%  u_1 in M  u1 < w
%  u_2 in M  u2 < w
%
The notion of which dimensioons $i$ have that \emph{any scheme in the model
has $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
%
\begin{definition}[Constrained dimensions]
Given an $n$-dimensional model $M$, the \emph{constrained dimensions}
of that model is a subset of $\{1, \ldots, n\}$ of 
the dimensions for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}

@@

The pairwise permutation $u \bowtie v$ takes two vectors and builds the matrix
where the rows are all possible $n$-vectors generated by
non-deterministic picking each $i^{th}$ entry from either the
$i^{th}$ entry in $u$ or the $i^{th}$ entry in $j$. For example:
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\ 
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
% \\
%\setlength{\arraycolsep}{0.5em}
%\left[\begin{array}{ccc}
%0 & 1 & 2 \\
%0 & 1 & 5 \\
%0 & 4 & 2 \\
%0 & 4 & 5 \\
%3 & 1 & 2 \\
%3 & 1 & 5 \\
%3 & 4 & 2 \\
%3 & 4 & 5
%\end{array}
%\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. Thus, we can
defined the pairwise permutation matrix as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
where $b$ is a logical matrix where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$.
%%
\end{definition}
